---
title: Descriptive analysis of the dataset of home sales in the city of Ames, Iowa for BAN 502 Project Phase 1.
author: "Prafull Pandey"
date: "`r Sys.Date()`"
output: word_document
---
# Exploratory/descriptive analysis of of home sales in the city of Ames, Iowa. The response variable in the dataset is “Above_Median”

```{r Sourcing Libraries}
library(tidyverse)
library(tidymodels)
library(GGally)
library(gridExtra) #used for a little fancy arranging of plots
library(car) #for the VIF function
library(glmnet)
library(skimr)
library(ggcorrplot) #create an alternative to ggcorr plots
library(MASS) #access to forward and backward selection algorithms
library(leaps) #best subset selection
library(lmtest) #for the dw test
library(splines) #for nonlinear fitting
library(car) #for calculating the variance inflation factor
```

## Read-in data. For this work we are using the dataset of home sales in the city of Ames, Iowa.

```{r}
ames=read_csv("ames_student-1.csv")
```


## Examined the ames data frame
```{r}
str(ames)
#glimpse(ames)
skim(ames)
```

## Examin summary of dataset and Perform mutation of response variable into numeric for quantitative variables corelation analysis and translate all character variables into factors.
```{r Data range selection to limit observations, mutate and translate all characters into factors}
summary(ames)
ames2= ames %>%
  mutate(Above_Median_Numeric = ifelse(Above_Median == "Yes", 1, 0)) # to estimate correlation with quantitative or continuous variables.
#ames2 = ames2 %>% filter(ames$Year_Built >= 1965) # Decided to consume full data

ames2 = ames2 %>% mutate_if(is.character, as_factor)  # translate all character variables to factor variables.

```

## Examin dataset after translation and mutation.


```{r Examin ames2 dataset}
summary(ames2)
skim(ames2)
```

## Visualize relationsship of all variables with response variable Above_Median to identify required variables for future model with expected positive relationship
```{r}
ggpairs(ames2, columns = c("MS_SubClass", "MS_Zoning", "Lot_Frontage", "Lot_Area","Above_Median"),cardinality_threshold = 50)
ggpairs(ames2, columns = c("Street", "Alley", "Lot_Shape", "Land_Contour","Above_Median"))
ggpairs(ames2, columns = c("Utilities", "Lot_Config", "Land_Slope", "Neighborhood","Above_Median"),cardinality_threshold = 50)
ggpairs(ames2, columns = c("Condition_1", "Condition_2", "Bldg_Type", "House_Style","Above_Median"))
ggpairs(ames2, columns = c("Overall_Qual", "Overall_Cond", "Year_Built", "Year_Remod_Add","Above_Median"))
ggpairs(ames2, columns = c("Roof_Style", "Roof_Matl", "Exterior_1st", "Exterior_2nd","Above_Median"),cardinality_threshold = 50)
ggpairs(ames2, columns = c("Mas_Vnr_Type", "Mas_Vnr_Area", "Exter_Qual", "Exter_Cond","Above_Median"))
ggpairs(ames2, columns = c("Foundation", "Bsmt_Qual", "Bsmt_Cond", "Bsmt_Exposure","Above_Median"))
ggpairs(ames2, columns = c("BsmtFin_Type_1", "BsmtFin_SF_1", "BsmtFin_Type_2", "BsmtFin_SF_2","Above_Median"))
ggpairs(ames2, columns = c("Bsmt_Unf_SF", "Total_Bsmt_SF", "Heating", "Heating_QC","Above_Median"))
ggpairs(ames2, columns = c("Central_Air", "Electrical", "First_Flr_SF", "Second_Flr_SF","Above_Median"))
ggpairs(ames2, columns = c("Low_Qual_Fin_SF", "Gr_Liv_Area", "Bsmt_Full_Bath", "Bsmt_Half_Bath","Above_Median"))
ggpairs(ames2, columns = c("Full_Bath", "Half_Bath", "Bedroom_AbvGr", "Kitchen_AbvGr","Above_Median"))
ggpairs(ames2, columns = c("Kitchen_Qual", "TotRms_AbvGrd", "Functional", "Fireplaces","Above_Median"))
ggpairs(ames2, columns = c("Fireplace_Qu", "Garage_Type", "Garage_Finish", "Garage_Cars","Above_Median"))
ggpairs(ames2, columns = c("Garage_Area", "Garage_Qual", "Garage_Cond", "Paved_Drive","Above_Median"))
ggpairs(ames2, columns = c("Wood_Deck_SF", "Open_Porch_SF", "Enclosed_Porch", "Three_season_porch","Above_Median"),cardinality_threshold = 251)
ggpairs(ames2, columns = c("Screen_Porch", "Pool_Area", "Pool_QC", "Fence","Above_Median"),cardinality_threshold = 251)
ggpairs(ames2, columns = c("Misc_Feature", "Misc_Val", "Mo_Sold", "Year_Sold","Above_Median"))
ggpairs(ames2, columns = c("Sale_Type", "Sale_Condition", "Longitude", "Latitude","Above_Median"))


```
### Analyze all variables and segrigating in two differfent groups to examin them as factor or quantitative variables.
```{r Segrigating and analyzing}

ames3= ames2[, sapply(ames2, is.factor)]
ames4= ames2[, !sapply(ames2, is.factor)]

```
### Analyze contineous quantitative variables with closer look on data patterns

```{r}
#colnames(ames4)
glimpse((ames4))
skim(ames4)
```


## We look at correlation for continuous quantitative variables. 
```{r}
ames2 %>% select_if(is.numeric) %>% dplyr::select("Lot_Frontage","Lot_Area", "Year_Built","Year_Remod_Add","Mas_Vnr_Area","BsmtFin_SF_1","BsmtFin_SF_2","Bsmt_Unf_SF","Above_Median_Numeric") %>% cor() %>%
ggcorrplot(hc.order = TRUE, type = "lower",    lab = TRUE)

ames2 %>% select_if(is.numeric) %>% dplyr::select("Total_Bsmt_SF","First_Flr_SF","Second_Flr_SF","Low_Qual_Fin_SF","Gr_Liv_Area","Bsmt_Full_Bath","Bsmt_Half_Bath","Full_Bath","Above_Median_Numeric") %>% cor() %>%
ggcorrplot(hc.order = TRUE, type = "lower",lab = TRUE)


ames2 %>% select_if(is.numeric) %>% dplyr::select("Half_Bath","Bedroom_AbvGr","Kitchen_AbvGr","TotRms_AbvGrd","Fireplaces","Garage_Cars","Garage_Area","Wood_Deck_SF","Above_Median_Numeric") %>% cor() %>%
ggcorrplot(hc.order = TRUE, type = "lower",lab = TRUE)

ames2 %>% select_if(is.numeric) %>% dplyr::select("Open_Porch_SF","Enclosed_Porch","Three_season_porch","Screen_Porch","Pool_Area","Misc_Val","Mo_Sold","Year_Sold","Above_Median_Numeric") %>% cor() %>%
ggcorrplot(hc.order = TRUE, type = "lower",lab = TRUE)

ames2 %>% select_if(is.numeric) %>% dplyr::select("Longitude","Latitude","Above_Median_Numeric") %>% cor() %>%
ggcorrplot(hc.order = TRUE, type = "lower",lab = TRUE)
 
```

### Analyze nominal variables with closer look on data patterns

```{r}
#colnames(ames3)
glimpse((ames3))
#labels(ames3)
skim(ames3)
```



## Identifying variables for final dataset
```{r Prepare final dataset}
homes_data=ames2 %>% dplyr::select("Above_Median", "Lot_Config","Overall_Qual","Neighborhood","Bldg_Type", "House_Style","Exterior_2nd","Bsmt_Exposure","BsmtFin_Type_1", "Heating_QC","Garage_Finish","Garage_Qual", "Year_Built", "Year_Remod_Add","Gr_Liv_Area" ,"Fireplaces","Garage_Cars","Garage_Area")
skim(homes_data)
glimpse(homes_data)
str(homes_data)
```

## Examin Corelation of nominal variables using focused plots since we had identified them during correlation plots in begining 
```{r Data visalization for Nominal group 1}

p1 = ggplot(homes_data, aes(x = Lot_Config, fill = Above_Median)) + geom_bar(position = "fill") + theme(axis.text.x = element_text(angle = 90, size = 8, vjust = 0.5, hjust=1))
p2 = ggplot(homes_data, aes(x = Overall_Qual, fill = Above_Median)) + geom_bar(position = "fill") + theme(axis.text.x = element_text(angle = 90, size = 8, vjust = 0.5, hjust=1))
p3 = ggplot(homes_data, aes(x = Bldg_Type, fill = Above_Median)) + geom_bar(position = "fill") + theme(axis.text.x = element_text(angle = 90, size = 8, vjust = 0.5, hjust=1))
p4 = ggplot(homes_data, aes(x = House_Style, fill = Above_Median)) + geom_bar(position = "fill") + theme(axis.text.x = element_text(angle = 90, size = 8, vjust = 0.5, hjust=1))
p5 = ggplot(homes_data, aes(x = Exterior_2nd, fill = Above_Median)) + geom_bar(position = "fill") + theme(axis.text.x = element_text(angle = 90, size = 8, vjust = 0.5, hjust=1))
p6 = ggplot(homes_data, aes(x = Bsmt_Exposure, fill = Above_Median)) + geom_bar(position = "fill") + theme(axis.text.x = element_text(angle = 90, size = 8, vjust = 0.5, hjust=1))
p7 = ggplot(homes_data, aes(x = BsmtFin_Type_1, fill = Above_Median)) + geom_bar(position = "fill") + theme(axis.text.x = element_text(angle = 90, size = 8, vjust = 0.5, hjust=1))
p8 = ggplot(homes_data, aes(x = Heating_QC, fill = Above_Median)) + geom_bar(position = "fill") + theme(axis.text.x = element_text(angle = 90, size = 8, vjust = 0.5, hjust=1))
p9 = ggplot(homes_data, aes(x = Garage_Finish, fill = Above_Median)) + geom_bar(position = "fill") + theme(axis.text.x = element_text(angle = 90, size = 8, vjust = 0.5, hjust=1))
p10 = ggplot(homes_data, aes(x = Garage_Qual, fill = Above_Median)) + geom_bar(position = "fill") + theme(axis.text.x = element_text(angle = 90, size = 8, vjust = 0.5, hjust=1))

grid.arrange(p1,p2,p3,p4, ncol = 2, heights=c(4,4))
grid.arrange(p5,p6,p6,p8, ncol = 2, heights=c(4,4))
grid.arrange(p9,p10, ncol = 2, heights=c(4,4))

```

```{r Data visalization for Nominal group 2}
ggplot(homes_data, aes(x = Neighborhood, fill = Above_Median)) + geom_bar(position = "fill") + theme(axis.text.x = element_text(angle = 60, size = 9, vjust = 0.6, hjust=0.6))
```




### Examing datapoint of Continuous variables in relation with response variable : Year_Built, Year_Remod_Add, Mas_Vnr_Area, First_Flr_SF, Fireplaces, Garage_Cars, Garage_Area	

```{r Data visalization for Continuous group 2}

p11 = ggplot(homes_data,aes(x=Above_Median,y=Year_Built)) + geom_boxplot() + geom_jitter(alpha = 0.15) + theme_bw() 
p12 = ggplot(homes_data,aes(x=Above_Median,y=Year_Remod_Add)) + geom_boxplot() + geom_jitter(alpha = 0.15) + theme_bw() 
p13 = ggplot(homes_data,aes(x=Above_Median,y=Gr_Liv_Area)) + geom_boxplot() + geom_jitter(alpha = 0.15) + theme_bw()
p14 = ggplot(homes_data,aes(x=Above_Median,y=Fireplaces)) + geom_boxplot() + geom_jitter(alpha = 0.15) + theme_bw() 
p15 = ggplot(homes_data,aes(x=Above_Median,y=Garage_Cars)) + geom_boxplot() + geom_jitter(alpha = 0.15) + theme_bw()
p16 = ggplot(homes_data,aes(x=Above_Median,y=Garage_Area)) + geom_boxplot() + geom_jitter(alpha = 0.15) + theme_bw() 

grid.arrange(p11,p12,p13,p14, ncol = 2)
grid.arrange(p15, p16, ncol = 2)

```

